For submission to the Journal of Integrated Pest Management

### 2021-04-08 ofv_fife_2021.Rmd
Gave up on the map concept, added a column in the table to indicate location where OFV was found

<!-- Several flat mite species, all from the genus *Brevipalpus* Donnadieu (Trombidiformes: Tenuipalpidae), are the only known vectors for the plant viruses known as cileviruses and dichorhaviruses. The *B. californicus* species group exclusively can transmit Orchid fleck dichorhavirus (OFV) in a persistent propagative manner. -->
OFV is the type species for the genus *Dichorhavirus* and infects more than fifty plant species belonging to the family Orchidaceae, Asparagaceae (Nolinoidaea), and Rutaceae (*Citrus*). During the summer of 2020, chlorotic ringspot symptoms were seen on Giant Lilyturf (*Liriope* spp., cv. ‘Gigantea’, Asparagaceae: Nolinoidaea) and leaf chlorosis was observed on leaves of *Aspidistra elatior* Blume (Asparagaceae: Nolinoidaea) in a landscape in Leon County, Florida. In both cases, the presence of OFV was confirmed using OFV specific conventional reverse transcription polymerase chain assay (RT-PCR) assay and Sanger sequencing. RT-PCR amplicons had a 98\% identity with the known OFV sequences available in the GenBank (Accession numbers: AB244418 and LC222630). The identification was also confirmed with the quantitative RT-PCR (RT-qPCR). Additional leaf samples were collected from other Nolinoidaea, including Aztec Grass (*Ophiopogon intermedius* cv. 'Argenteomarginatus' D. Don, Asparagaceae: Nolinoidaea), Mondo Grass *O. japonicus* and Lilyturf *L. muscari*, with ringspot symptoms were observed in Leon and Alachua counties. Identification of partial genome sequence confirmed the presence of both the orchid strains (OFV-Orc1 and OFV-Orc2) in Florida. Three mite species were identified from OFV-infected plants: *Brevipalpus californicus* sensu lato, *B. obovatus* Donnadieu and *B. confusus* Banks. At least one of these species is presumably responsible for OFV transmission. Florida has various *Brevipalpus*susceptible native and introduced plant species in the landscape. In this study we are reporting three new hosts from the family Asperagaceae from multiple locations. Our data suggested that the OFV is widely distributed in Florida and might be a potential threat for *Liriope* spp., *Ophiopogon* spp. and *Aspidistra elatior* which are commonly used in landscaping in the southeastern United States, if not controlled. As both the orchid strains of OFV are known to infect citrus and cause citrus leprosis disease, a survey in citrus growing regions of Florida is essential, emphasizing plants within the families Orchidaceae, Rutaceae and Asparagaceae.

@@

OFV-infected plants exhibit various symptoms depending on the infected plant species as well as the strain of the OFV associated with the infection [@Kubo2009b], but symptoms typically appear as chlorotic flecks, which ultimately coalesce into larger spots or ringspot patterns (`r citef("fig_1")`, `r citef("fig_2")`).
<!-- These lesions are thought to be associated with the feeding sites of the mite vector [@cite?]. Also, is it possibly a sign of RNAi plant defenses? See Hull Plant Virology -->

@@

Subsequent surveys of plants belonging to the subfamily Nolinoidaea in Florida have revealed more sites with Asparagaceae putatively infected with OFV in Leon and Alachua counties (`r citet("tab_1")`, `r citef("fig_3")`). Mites were collected from symptomatic plants in Leon county and observed with phase contrast microscopy. Tenuipalpid mites (flat mites or false spider mites) were commonly found in abundance on the Asparagaceae which tested positive for OFV. These flat mites were originally identified as *Brevipalpus californicus* (Banks) sensu lato and later confirmed by the FDACS via Differential Interference Contrast (DIC) microscopy. The *Brevipalpus* mite species complex is known to contain cryptic species [@Childers2011] that require advanced microscopy techniques, such as cryo scanning electron microscopy (Cryo-SEM) for species identification [@Skoracka2015; @Leon2010; @Beard2015]. Additional mite samples were collected from the original OFV detection site, and examined under Cryo-SEM (`r citef("fig_4")`). The determinations approved prior identifications of *B. californicus* s.l. but revealed the presence of two other species *B. obovatus* and *B. confusus* (`r citef("fig_5")`).

@@

### Conclusion
The dichorhavirus that infects citrus in Hawaii, Mexico, Colombia, and South Africa are identical to the OFV in gene order, content, and the genome sequence. According to the International committee on Taxonomy of Viruses (ICTV) classification, OFV consist of two orchid strains (OFV-Orc1 and OFV-Orc2) and two citrus strains (OFV-Cit1 and OFV-Cit2). Both the orchid strains of OFV infects citrus [@Roy2020], but none of the citrus strains have been reported from any orchid species. Detecting OFV in Florida represents a concern for horticulturists who grow orchids, *Liriope*, *Ophiopogon*, or other susceptible Asparagaceae species which are commonly used in landscaping. Florida is also home to a plethora of native and naturalized orchid species, many of which are threatened, including cultivating *Vanilla* in southern Florida [@Chambers2019] and the famous Ghost Orchid, [*Dendrophylax lindenii* (Lindl.) Benth. ex Rolfe]. Citrus leprosis was present in Florida during the 1860’s and eradicated in the mid-1960s. In fact, @Kitajima2011 found that the Citrus Leprosis virus (CiLV) which previously affected Florida citrus was a nuclear type of citrus leprosis closely related to OFV strains [@Roy2020; @Hartung2015]. Association of a distant relative of OFV named Citrus leprosis dichorhavirus-N0 (CiLV-N0) was confirmed in relation to the leprosis disease outbreak in Florida [@Hartung2015]. The recent detection of OFV-Orc1 in South Africa [@Cook2019] in *C. sinensis* (Navel and Valencia orange) and OFV-Orc2 in Hawaii [@Velarde2021] in *C. reticulata* (mandarin) and *C. jambhiri* (rough lemon) associated with leprosis-like symptoms highlights the threat of different strains of OFV on citrus; which will be a definite concern to the multi-billion dollar citrus industry. *B. californicus* and *B. yothersi* are both known vectors of Dichorhaviruses (OFV) and Cileviruses (Citrus Leprosis) [@Knorr1968; @Kondo2003; @Beltran-Beltran2020] and *B. obovatus* is a suspected vector as well [@Childers2003]. All three mite species/complexes are present in Florida [@Childers2003; @Akyazi2017] (`r citef("fig_5")`). Therefore, it is critical to identify the vector of OFVs in Florida and monitor its spread to determine the risk this virus represents for the native plants, agriculture and the ornamental/landscaping industries of Florida and the surrounding regions.

@@

|Scientific Name|Common Names|Symptoms Observed|
|:-----------|:--------------|:----------------|
|*Liriope muscari* Bailey| Lilyturf, Orchardgrass, Monkeygrass| Ringspots, Chlorotic Flecking, Necrotic Lesions|
|*Liriope gigantea*\* Hume| Giant Lilyturf| Ringspots, Chlorotic Flecking, Necrotic Lesions|
|*Ophiopogon japonicus* Ker Gawl.| Dwarf Lilyturf, Mondo Grass, Snake's beard| Ringspots, Chlorotic Flecking, Necrotic Lesions|
|*Ophiopogon intermedius*\*\* Don| Aztec Grass, 'Argenteomarginatus' | Ringspots, Chlorotic Flecking, Necrotic Lesions|
|*Aspidistra elatior* Blume| Cast Iron Plant, Bar-room Plant | Chlorosis, Chlorotic Flecking, Necrotic Lesions|
`r tab_nums(name = "tab_1", caption = "List of plants with symptoms of Orchid fleck dichorhavirus found in northern Florida. * *L. gigantea* have been traditionally classified as seperate from *L. muscari* by @Broussard2007 and @Fantz2015, although this distinction has been challenged by @Wang2014 and @Masiero2020. * * *O. intermedius* is sometimes misclassified as *Liriope muscari* ‘Variegated Evergreen Giant’ [@Fantz2008b, @Fantz2009] or ‘Grandiflora White’ [@Fantz2009].")`

@@

`r fig_nums(name = "fig_3", caption = "Map of OFV cases in Florida")`

`r fig_nums(name = "fig_4", caption = "Cryo-SEM images of *Brevipalpus californicus* sensu lato displaying various characters used for identification [@Baker1987; @Beard2012] (a) Dorsum  (b) Lateral view (c) Venter (d) Close up of distal end of leg 2, with arrows indicating paired solenidia, characteristic of the genus *Brevipalpus* (e) Enlargement of the microplates of the mite cerotegument (f) Dorsal view of the distal portion of mite abdomen (g) Dorsal view of the mite rostrum (h) Ventral view of mite rostrum, observe 3 distal setae.")`

`r fig_nums(name = "fig_5", caption = "Florida is home to other common pest species of *Brevipalpus* which are potential vectors of *Orchid fleck dichorhavirus*: (a) *B. yothersi*, dorsal (b) *B. yothersi*, lateral (c) *B. obovatus*, dorsal.")`

@@

```{r, fig_3, warning=FALSE}
grid::grid.raster(tiff::readTIFF("figure/map_ofv_fl.tif"))
```

```{r, fig_4, warning=FALSE}
grid::grid.raster(tiff::readTIFF("figure/b_cali.tif"))
```

```{r, fig_5, warning=FALSE}
grid::grid.raster(tiff::readTIFF("figure/other_mites.tif"))
```

>>>>

Several flat mite species, all from the genus *Brevipalpus* Donnadieu (Trombidiformes: Tenuipalpidae), are the only known vectors for the plant viruses known as cileviruses and dichorhaviruses. The *B. californicus* species group exclusively can transmit Orchid fleck dichorhavirus (OFV) in a persistent propagative manner.
OFV is the type species for the genus *Dichorhavirus* and infects more than fifty plant species belonging to the family Orchidaceae, Asparagaceae (Nolinoidaea), and Rutaceae (*Citrus*). During the summer of 2020, chlorotic ringspot symptoms were seen on Giant Lilyturf (*Liriope* spp., cv. ‘Gigantea’, Asparagaceae: Nolinoidaea) and leaf chlorosis was observed on leaves of *Aspidistra elatior* Blume (Asparagaceae: Nolinoidaea) in a landscape in Leon County, Florida. In both cases, the presence of OFV was confirmed using OFV specific conventional reverse transcription polymerase chain assay (RT-PCR) assay and Sanger sequencing. RT-PCR amplicons had a 98\% identity with the known OFV sequences available in the GenBank (Accession numbers: AB244418 and LC222630). The identification was also confirmed with the quantitative RT-PCR (RT-qPCR). Additional leaf samples were collected from other Nolinoidaea, including Aztec Grass (*Ophiopogon intermedius* cv. 'Argenteomarginatus' D. Don, Asparagaceae: Nolinoidaea), Mondo Grass *O. japonicus* and Lilyturf *L. muscari*, with ringspot symptoms were observed in Leon and Alachua counties. Identification of partial genome sequence confirmed the presence of both the orchid strains (OFV-Orc1 and OFV-Orc2) in Florida. Three mite species were identified from OFV-infected plants: *Brevipalpus californicus* sensu lato, *B. obovatus* Donnadieu and *B. confusus* Banks. At least one of these species is presumably responsible for OFV transmission. Florida has various *Brevipalpus*susceptible native and introduced plant species in the landscape. In this study we are reporting three new hosts from the family Asperagaceae from multiple locations. Our data suggested that the OFV is widely distributed in Florida and might be a potential threat for *Liriope* spp., *Ophiopogon* spp. and *Aspidistra elatior* which are commonly used in landscaping in the southeastern United States, if not controlled. As both the orchid strains of OFV are known to infect citrus and cause citrus leprosis disease, a survey in citrus growing regions of Florida is essential, emphasizing plants within the families Orchidaceae, Rutaceae and Asparagaceae.

@@

OFV-infected plants exhibit various symptoms depending on the infected plant species as well as the strain of the OFV associated with the infection [@Kubo2009b], but symptoms typically appear as chlorotic flecks, which ultimately coalesce into larger spots or ringspot patterns (`r citef("fig_1")`, `r citef("fig_2")`).

@@

Subsequent surveys of plants belonging to the subfamily Nolinoidaea in Florida have revealed more sites with Asparagaceae putatively infected with OFV in Leon and Alachua counties (`r citet("tab_1")`). Mites were collected from symptomatic plants in Leon county and observed with phase contrast microscopy. Tenuipalpid mites (flat mites or false spider mites) were commonly found in abundance on the Asparagaceae which tested positive for OFV. These flat mites were originally identified as *Brevipalpus californicus* (Banks) sensu lato and later confirmed by the FDACS via Differential Interference Contrast (DIC) microscopy. The *Brevipalpus* mite species complex is known to contain cryptic species [@Childers2011] that require advanced microscopy techniques, such as cryo scanning electron microscopy (Cryo-SEM) for species identification [@Skoracka2015; @Leon2010; @Beard2015]. Additional mite samples were collected from the original OFV detection site, and examined under Cryo-SEM (`r citef("fig_3")`). The determinations approved prior identifications of *B. californicus* s.l. but revealed the presence of two other species *B. obovatus* and *B. confusus* (`r citef("fig_4")`).

@@

### Conclusion
The dichorhavirus that infects citrus in Hawaii, Mexico, Colombia, and South Africa are identical to the OFV in gene order, content, and the genome sequence. According to the International committee on Taxonomy of Viruses (ICTV) classification, OFV consist of two orchid strains (OFV-Orc1 and OFV-Orc2) and two citrus strains (OFV-Cit1 and OFV-Cit2). Both the orchid strains of OFV infects citrus [@Roy2020], but none of the citrus strains have been reported from any orchid species. Detecting OFV in Florida represents a concern for horticulturists who grow orchids, *Liriope*, *Ophiopogon*, or other susceptible Asparagaceae species which are commonly used in landscaping. Florida is also home to a plethora of native and naturalized orchid species, many of which are threatened, including cultivating *Vanilla* in southern Florida [@Chambers2019] and the famous Ghost Orchid, [*Dendrophylax lindenii* (Lindl.) Benth. ex Rolfe]. Citrus leprosis was present in Florida during the 1860’s and eradicated in the mid-1960s. In fact, @Kitajima2011 found that the Citrus Leprosis virus (CiLV) which previously affected Florida citrus was a nuclear type of citrus leprosis closely related to OFV strains [@Roy2020; @Hartung2015]. Association of a distant relative of OFV named Citrus leprosis dichorhavirus-N0 (CiLV-N0) was confirmed in relation to the leprosis disease outbreak in Florida [@Hartung2015]. The recent detection of OFV-Orc1 in South Africa [@Cook2019] in *C. sinensis* (Navel and Valencia orange) and OFV-Orc2 in Hawaii [@Velarde2021] in *C. reticulata* (mandarin) and *C. jambhiri* (rough lemon) associated with leprosis-like symptoms highlights the threat of different strains of OFV on citrus; which will be a definite concern to the multi-billion dollar citrus industry. *B. californicus* and *B. yothersi* are both known vectors of Dichorhaviruses (OFV) and Cileviruses (Citrus Leprosis) [@Knorr1968; @Kondo2003; @Beltran-Beltran2020] and *B. obovatus* is a suspected vector as well [@Childers2003]. All three mite species/complexes are present in Florida [@Childers2003; @Akyazi2017] (`r citef("fig_4")`). Therefore, it is critical to identify the vector of OFVs in Florida and monitor its spread to determine the risk this virus represents for the native plants, agriculture and the ornamental/landscaping industries of Florida and the surrounding regions.

@@

|Scientific Name|Common Names|Symptoms Observed|County|
|:-----------|:--------------|:----------------|:-----|
|*Liriope muscari* Bailey| Lilyturf, Orchardgrass, Monkeygrass| Ringspots, Chlorotic Flecking, Necrotic Lesions|Alachua & Leon|
|*Liriope gigantea*\* Hume| Giant Lilyturf| Ringspots, Chlorotic Flecking, Necrotic Lesions|Alachua & Leon|
|*Ophiopogon japonicus* Ker Gawl.| Dwarf Lilyturf, Mondo Grass, Snake's beard| Ringspots, Chlorotic Flecking, Necrotic Lesions|Leon|
|*Ophiopogon intermedius*\*\* Don| Aztec Grass, 'Argenteomarginatus' | Ringspots, Chlorotic Flecking, Necrotic Lesions|Alachua & Leon|
|*Aspidistra elatior* Blume| Cast Iron Plant, Bar-room Plant | Chlorosis, Chlorotic Flecking, Necrotic Lesions|Leon|
`r tab_nums(name = "tab_1", caption = "List of plants with symptoms of Orchid fleck dichorhavirus found in northern Florida. * *L. gigantea* have been traditionally classified as seperate from *L. muscari* by @Broussard2007 and @Fantz2015, although this distinction has been challenged by @Wang2014 and @Masiero2020. * * *O. intermedius* is sometimes misclassified as *Liriope muscari* ‘Variegated Evergreen Giant’ [@Fantz2008b, @Fantz2009] or ‘Grandiflora White’ [@Fantz2009].")`

@@

`r fig_nums(name = "fig_3", caption = "Cryo-SEM images of *Brevipalpus californicus* sensu lato displaying various characters used for identification [@Baker1987; @Beard2012] (a) Dorsum  (b) Lateral view (c) Venter (d) Close up of distal end of leg 2, with arrows indicating paired solenidia, characteristic of the genus *Brevipalpus* (e) Enlargement of the microplates of the mite cerotegument (f) Dorsal view of the distal portion of mite abdomen (g) Dorsal view of the mite rostrum (h) Ventral view of mite rostrum, observe 3 distal setae.")`

`r fig_nums(name = "fig_4", caption = "Florida is home to other common pest species of *Brevipalpus* which are potential vectors of *Orchid fleck dichorhavirus*: (a) *B. yothersi*, dorsal (b) *B. yothersi*, lateral (c) *B. obovatus*, dorsal.")`

@@

```{r, fig_3, warning=FALSE}
grid::grid.raster(tiff::readTIFF("figure/b_cali.tif"))
```

```{r, fig_4, warning=FALSE}
grid::grid.raster(tiff::readTIFF("figure/other_mites.tif"))
```
